---
title: "Reproducible Research"
subtitle: "Project One"
author: 
- "Submitted by: Pete Ostergren"
- "----------------------------------------"
date: "11/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## 1. Load data
Unzip data to obtain a csv file.

```{r unzip, cache=TRUE}
unzip("data/repdata_data_activity.zip", exdir = "data")
activity <- read.csv("data/activity.csv", stringsAsFactors=FALSE)
```

## 2. Histogram
Use dplyr.

```{r daily_total,message=FALSE}
require(dplyr)
total_day <- activity %>% group_by(date) %>%summarise(total_steps=sum(steps,na.rm=TRUE),na=mean(is.na(steps)))

total_day <- filter(total_day, na < 1)
hist(total_day$total_steps,col="orange",breaks=20,main="Total steps per day",xlab="Steps per day")
abline(v=median(total_day$total_steps),lty=3, lwd=2, col="black")
legend(legend="median","topright",lty=3,lwd=2,bty = "n")
```
## 3. Mean and median number of steps taken each day

```{r summary}
mean_steps <- mean(total_day$total_steps,na.rm=TRUE)
median_steps <- median(total_day$total_steps,na.rm=TRUE)
```
## 4. Time series plot of the average number of steps taken

```{r daily, fig.width=12}
library(dplyr,quietly = TRUE)
daily_patterns <- activity %>% group_by(interval) %>% summarise(average=mean(steps,na.rm=TRUE))
plot(x = 1:nrow(daily_patterns),y = daily_patterns$average,type = "l",
     col = "red", xaxt = "n",xlab="Intervals", 
     ylab = "Average for given interval across all days")
axis(1,labels=daily_patterns$interval[seq(1,288,12)],
     at = seq_along(daily_patterns$interval)[seq(1,288,12)])
```


## 5. The 5-minute interval containing the max avg number of steps

```{r daily2}
max <- filter(daily_patterns,average==max(average))
```

Interval **"`r max$interval`"** contains on average the maximum number of steps (**`r round(max$average,2)` steps**).


## 6. Impute missing data

```{r na_imputing, cache=TRUE}
without_NAs <- numeric(nrow(activity))
for (i in 1:nrow(activity))
{
        if (is.na(activity[i,"steps"])==TRUE)
            {
                    without_NAs[i]<-filter(daily_patterns,interval==activity[i,"interval"]) %>% select(average)
            } 
        else
            {
                    without_NAs[i]<-activity[i,"steps"]
            }
                    
}
activity_without_NAs<-mutate(activity,steps_no_NAs=without_NAs)
head(activity_without_NAs)
```

## 7. Histogram: Total number of steps taken each day with missing values imputed

```{r histogram_no_NAs}
total_day_noNAs <- activity_without_NAs %>% mutate(steps_no_NAs=as.numeric(steps_no_NAs)) %>% group_by(date) %>% summarise(total_steps=sum(steps_no_NAs))
hist(total_day_noNAs$total_steps,col="blue",breaks=20,main="Total steps per day",xlab="Steps per day")
abline(v=median(total_day$total_steps),lty=3, lwd=2, col="black")
legend(legend="median","topright",lty=3,lwd=2,bty = "n")
```

## 8. Panel plot comparing the average number of steps taken per 5-minute interval with weekday vs weekend

```{r weekday}
library(lubridate)
is_weekday <-function(date){
        if(wday(date)%in%c(1,7)) result<-"weekend"
        else
                result<-"weekday"
        result
}
activity_without_NAs <- mutate(activity_without_NAs,date=ymd(date)) %>% mutate(day=sapply(date,is_weekday))
table(activity_without_NAs$day)
```
```{r weekend_comparison, cache=TRUE, fig.width=10}
library(ggplot2)
daily_patterns <- activity_without_NAs %>% mutate(day=factor(day,levels=c("weekend","weekday")),steps_no_NAs=as.numeric(steps_no_NAs)) %>% group_by(interval,day) %>% summarise(average=mean(steps_no_NAs))
qplot(interval,average,data=daily_patterns,geom="line",facets=day~.)
```